Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-04-07 WSSSPE2_report/WSSSPE2_review.tex
Section4: raw summary.

Progress in scientific research is dependent on the quality and accessibility of
software at all levels. Hence it is critical to address challenges related to
the development, deployment, maintenance, and overall sustainability of reusable
software as well as education around software practices. These challenges can be
technological, policy based, organizational, and educational, and are of
interest to developers (the software community), users (science disciplines),
software-engineering researchers, and researchers studying the conduct of
science (science of team science, science of organizations, science of science
and innovation policy, and social science communities). The WSSSPE1 workshop
engaged the broad scientific community to identify challenges and best practices
in areas of interest to creating sustainable scientific software. WSSSPE2
invited the community to propose and discuss specific mechanisms to move towards
an imagined future practice for software development and usage in science and
engineering. The workshop included multiple mechanisms for participation,
encouraged team building around solutions, and identified risky solutions with
potentially transformative outcomes. It strongly encouraged participation of
early-career scientists, postdoctoral researchers, and graduate students.

@@

Neil Chue Hong is the Director of the Software Sustainability Institute (SSI) in the 
United Kingdom. The SSI was founded to support the UK?s research software 
community by cultivating better, more sustainable, research software to enable 
world-class research. Chue Hong's keynote\footnote{Slides available at: http://dx.doi.org/10.6084/m9.figshare.1243288}
 started by making the point of that the use of 
- and reliance on - software is pervasive in all areas of world-leading research, 
showing examples from disciplines as diverse as the humanities and high-energy 
physics, and quoting Kersten Kleese Van Dam of the Pacific Northwestern 
National Laboratory via a petition campaign\footnote{https://www.change.org/p/everyone-in-the-research-community-we-must-accept-that-software-is-fundamental-to-research-or-we-will-lose-our-ability-to-make-groundbreaking-discoveries}} 
at change.org, ``\emph{Today there are very few science areas left 
which do not rely on IT and thus software for the majority of their research work. 
More importantly key scientific advances in experimental and observational science 
would have been impossible without better software.}'' He also cited Daniel Katz, 
Software Infrastructure for Sustained Innovation(SI2) Program Director of the 
National Science Foundation, ``\emph{Scientific discovery and innovation are 
advancing along fundamentally new pathways opened by development of 
increasingly sophisticated software. Software is an integral enabler of 
computation, experiment and theory, and directly responsible for increased 
scientific productivity and enhancement of researchers' capabilities.}''

@@

%\item Robert R. Downs: Community Recommendations for Improving Sustainable
%Scientific Software Practices

>>>>

Progress in scientific research is dependent on the quality and accessibility of
software at all levels. Hence it is critical to address challenges related to
the development, deployment, maintenance, and overall sustainability of reusable
software as well as education around software practices. These challenges can be
technological, policy based, organizational, and educational, and are of
interest to developers (the software community), users (science disciplines),
software-engineering researchers, and researchers studying the conduct of
science (science of team science, science of organizations, science of science
and innovation policy, and social science communities). The WSSSPE1 workshop
engaged the broad scientific community to identify challenges and best practices
in areas of interest to creating sustainable scientific software. WSSSPE2
invited the community to propose and discuss specific mechanisms to move towards
an imagined future practice for software development and usage in science and
engineering. The workshop included multiple mechanisms for participation,
encouraged team building around solutions, and identified risky solutions with
potentially transformative outcomes. It strongly encouraged participation of
early-career scientists, postdoctoral researchers, and graduate students.
\todo{record how/number the travel awards were created and recipients selected, etc?}

@@

Neil Chue Hong is the Director of the Software Sustainability Institute (SSI) in the 
United Kingdom. The SSI was founded to support the UK?s research software 
community by cultivating better, more sustainable, research software to enable 
world-class research. Chue Hong's keynote\footnote{Slides available at: http://dx.doi.org/10.6084/m9.figshare.1243288}
 started by making the point of that the use of 
- and reliance on - software is pervasive in all areas of world-leading research, 
showing examples from disciplines as diverse as the humanities and high-energy 
physics, and quoting Kersten Kleese Van Dam of the Pacific Northwestern 
National Laboratory via a petition campaign\footnote{https://www.change.org/p/everyone-in-the-research-community-we-must-accept-that-software-is-fundamental-to-research-or-we-will-lose-our-ability-to-make-groundbreaking-discoveries}
at change.org, ``\emph{Today there are very few science areas left 
which do not rely on IT and thus software for the majority of their research work. 
More importantly key scientific advances in experimental and observational science 
would have been impossible without better software.}'' He also cited Daniel Katz, 
Software Infrastructure for Sustained Innovation(SI2) Program Director of the 
National Science Foundation, ``\emph{Scientific discovery and innovation are 
advancing along fundamentally new pathways opened by development of 
increasingly sophisticated software. Software is an integral enabler of 
computation, experiment and theory, and directly responsible for increased 
scientific productivity and enhancement of researchers' capabilities.}''

@@

\item \textbf{Robert R. Downs: Community Recommendations for Improving Sustainable
Scientific Software Practices.}
Robert R. Downs, of the Columbia University Center for International Earth
Science Information Network (CIESIN), presented Community Recommendations for
Improving Sustainable Scientific Software Practices (2014), and described a
focus group study that was conducted with other members of the Science Software
Cluster of the Federation of Earth Science Information Partners (ESIP),
including W. Christopher Lenhardt, Erin Robinson, Ethan Davis, and Nicholas
Weber. For the study, almost 300 attendees of the 2014 Summer ESIP Meeting were
invited to participate in simultaneous roundtable discussions, on the
sustainability of science software. Over two-thirds of the roundtable focus
groups provided responses to a semi-structured questionnaire that contained
three sets of questions. Initial analysis of the participants' responses to the
set of questions eliciting recommendations for near-term actions of the
community to improve sustainable software practices revealed several
suggestions. The recommendations for improving scientific software
sustainability included improving community engagement and collaborative
activities, increasing understanding and awareness, and creating incentives to
motivate sustainable science software practices. Based on the suggestions
offered by the focus group respondents for near-term actions, the ESIP Science
Software Cluster plans to engage the community to achieve progress in the near
future on the recommended activities for improving sustainable scientific
software practices.

Reference

Downs RR, Lenhardt WC, Robinson E, Davis E, Weber N. 2014. Community
Recommendations for Improving Sustainable Scientific Software Practices. 2nd
Workshop on Sustainable Software for Science: Practice and Experiences
(WSSSPE2), New Orleans, Louisiana, 16 November 2014.
http://wssspe.researchcomputing.org.uk/wp-content/uploads/2014/11/Downs.pdf and
http://dx.doi.org/10.7916/D8Q52NBC
\choinote{Choi to revise.}
